### 2015-10-08 docs/AIAA 2015/high_eikon.tex
no message

% !TEX root = main.tex
\subsection{The Eikonal Equation -- Cost-Minimizing Path}