### 2015-12-16 docs/AIAA 2015/main.tex

\documentclass[letterpaper]{AIAA}

\newcommand{\pos}{p}
\newcommand{\vel}{v}
\newcommand{\hw}{\mathbb{H}}
\newcommand{\hwd}{\hat{d}}
\newcommand{\cost}[1]{c_\text{#1}}
\newcommand{\hpath}{\mathcal{P}}
\newcommand{\wpt}{\mathcal{W}}
\newcommand{\cmap}{c}
\newcommand{\ccost}{C}
\newcommand{\ppath}{\mathbb{P}}
\newcommand{\ocost}{V}

\title{Guaranteeing Safety and Liveness of Unmanned Aerial Vehicle Platoons on Air Highways}

\author{Mo Chen\footnote{Insert Job Title, Department Name, Address/Mail Stop, and AIAA Member Grade (if any) for first author.}}
\affiliation{Business or Academic Affiliation 1, City, State, Zip Code}
\author{Third C. Author\footnote{Insert Job Title, Department Name, Address/Mail Stop, and AIAA Member Grade (if any) for third author.}}
\affiliation{Business or Academic Affiliation 2, City, Province, Zip Code, Country}
\author{Fourth D. Author\footnote{Insert Job Title, Department Name, Address/Mail Stop, and AIAA Member Grade (if any) for fourth author (etc.).}}
\affiliation{Business or Academic Affiliation 2, City, State, Zip Code}

\begin{abstract}
\end{abstract}

\maketitle

\section*{Nomenclature}
(Nomenclature entries should have the units identified)\\
\noindent\begin{tabular}{@{}lcl@{}}
\textit{x} &=& System state \\
\end{tabular} \\

\input{introduction.tex}
\input{high_form.tex}
\input{high_eikon.tex}
\input{high_waypoints.tex}
\input{high_sim.tex}
\input{plat_form.tex}
\input{plat_reach.tex}
\input{plat_intra.tex}
\input{plat_inter.tex}
\input{plat_safety.tex}
\input{plat_sim.tex}
\input{conclusions.tex}

\section*{Appendix}

\section*{Acknowledgments}
This work is supported in part by NSF under CPS:ActionWebs (CNS-0931843) and CPS:FORCES (CNS1239166), by NASA under grants NNX12AR18A and UCSCMCA-14-022 (UARC), by ONR under grants N00014-12-1-0609, N000141310341 (Embedded Humans MURI), and MIT\_5710002646 (SMARTS MURI), and by AFOSR under grants UPenn-FA9550-10-1-0567 (CHASE MURI) and the SURE project.

\section*{References}

\begin{thebibliography}{}
\bibitem{1} J. A. Sethian , ``A fast marching level set method for monotonically advancing fronts,'' \textit{Proceedings of the National Academy of Sciences of the United States of America}, Vol. 93, 1996, pp. 1591-1595

\end{thebibliography}
\end{document}