### 2016-02-17 docs/AIAA 2016/high_waypoints.tex

% !TEX root = main.tex
\subsection{From Paths to Waypoints}
Each of the cost-minimizing paths $\ppath_i$ computed from the solution to the Eikonal equation consists of a continuous set of points. Each path $\ppath_i$ is an approximation to the sequence of highways $\hws_{N_i}^i = \{\hw^i_j\}_{i=1,j=1}^{i=M,j=N_i}$ defined in \eqref{eq:ahpp}, but now indexed by the corresponding path index $i$. The

For each path $\ppath_i$, we would like to sparsify the points on the path to obtain a collection of waypoints, $\wpt_{i,j}, j = 1,\ldots, N_i+1$, which are the end points of the highways:

\begin{equation}
\begin{aligned}
\hw^i_j(0) &= \wpt_{i,j}, \\
\hw^i_j(1) &= \wpt_{i,j+1}, \\
j &= 1,\ldots,N_i
\end{aligned}
\end{equation}

To sparsify the continuous set of points on each path, we start at the destination point and note the path's heading. We add to the collection of waypoints the first point on the path at which the heading changes by some threshold $\theta_C$, and repeat this process is repeated along the entire path.

If there is a large change in heading within a small section of the cost-minimizing path, then the collection of point may contain many points which are close together. In addition, there may be multiple paths that are very close to each other (in fact, this behavior is desirable), which may contribute to cluttering the airspace with too many waypoints. We propose to sparsify the waypoints by clustering the points. After clustering, we replace each cluster of points with a single point located at the centroid of the cluster.

To the collection of points resulting from the above process generates we add the origin and destination points to the collection. Repeating the entire process for every path, we waypoints for all the cost-minimizing paths under consideration. Figure \ref{fig:hw_ill} summarizes the entire air highway placement process.