### 2016-02-25 docs/CDC 2016 multi-agent/introduction.tex
Renamed MIP paper folder

% !TEX root = MIP.tex
\section{Introduction}
- motivation
-- UAVs for civil purposes
-- difficult problem

- previous methods for dealing with multi-agent systems
-- references
-- limitations (especially with respect to HJ formulation)
--- linear/simple dynamics
--- non-cooperative settings
--- applications that don't directly address the problem

- HJ background
-- references
-- normal limitations

- HJ for multi-UAV systems
-- assumed structure (platooning, SPP)

- summary