### 2016-02-29 docs/CDC 2016 multi-agent/introduction.tex
Updated MIP paper introduction

% !TEX root = MIP.tex
\section{Introduction}
- motivation
-- UAVs for civil purposes \cite{Tice91, Debusk10, FAA13, AUVSI16, Amazon16, BBC16, NASA16}. For example, package delivery, aerial surveillance, disaster response, ...
-- difficult problem

- previous methods for dealing with multi-agent systems
-- limitations (especially with respect to HJ formulation)

In \cite{Tanimoto78}, the author discuss a class of three-player differential game in which two of the players oppose the third player in order to neutralize the third player's attempts to alter their joint cost function.

\cite{Lin13} considers multiple agent games in which communication follows a graphical structure.

In \cite{Su14}, the authors propose a control law derived from a differential game that models the protection of a target from a homing missile.

Most recently, \cite{Fisac15} uses time-dependent Hamilton-Jacobi reachability analysis to approximate the solution to a pursuit-evasion game in which a third player, the defender, attempts to capture the pursuer.

However, all of the above makes assumptions such as linear dynamics and the assumption of agent roles, and describes non-cooperative situations, making them insufficient for addressing the multi-vehicle collision avoidance problem

- HJ background
-- references
-- theory \cite{Mitchell05}
-- success in applications \cite{Ding08, Chen14}
-- normal limitations

- HJ for multi-UAV systems
-- \cite{Chen15, Chen15b, Chen16}
-- assumed structure (platooning, SPP)

- summary